 Casting multiple shadows: interactive data visualisation with tours and embeddings

### 2020-02-06 paper.Rmd
remove whitespace
finished first draft of intro!

A complimentary approach for visualizing structure in high dimensional data
is the tour. A tour is a sequence of projections of a high dimensional dataset
onto a low-dimensional orthonormal basis matrix, that is represented as a dynamic visualization. The sequence of generated bases are interpolated to form the tour path, allowing a user to explore the subspace of projections. A
grand tour corresponds to choosing new bases at random, and can give an
overview of the structure in the data. Instead of picking projections at random, a guided tour can be used to generate a sequence 'interesting' projections
as quantified by an index function. The tour has been used previously 
by @Wickham2015-cx as tool for exploring statistical model fits.

## Interactive graphics

@@

There is no one-size fits all: finding an appropriate embedding for a given
dataset is a difficult and somewhat poorly defined problem. For non-linear methods, there are a lot of parameters to explore that can have an effeft on the resulting visualisation and interpretation. We propose a more pragmatic workflow inspired by (wickham blindfold paper) by incorporating
interactive graphics and tours with embeddings that allows users to see a
global overview of their high dimensional data and assists them with cluster
orientation tasks. This workflow is incorporated into an R package
called `liminal` (Available: https://github.com/sa-lee/liminal).

>>>>

@@

A complimentary approach for visualizing structure in high dimensional data
is the tour. A tour is a sequence of projections of a high dimensional dataset
onto a low-dimensional orthonormal basis matrix, that is represented as a dynamic visualization. The sequence of generated bases are interpolated to form the tour path, allowing a user to explore the subspace of projections. A
grand tour corresponds to choosing new bases at random, and can give an
overview of the structure in the data. Instead of picking projections at random, a guided tour can be used to generate a sequence 'interesting' projections
as quantified by an index function. Given the dynamic nature of the tour, user interaction is important for controlling and exploring the visualisation: the tour has been used previously by @Wickham2015-cx as tool for exploring statistical model fits and by @Buja1996-fk for exploring factorial experimental
designs.

While there has been much work on the algorithmic details of the aformentioned
embedding methods, there has been relatively few tools designed to assist
users to interact with these techniques and assist them in making comparsions
between embeddings and performing the aforementioned cluster orientation tasks. 
Several interactive interfaces have been proposed for evaluating or using
embedding techniques: the Sleepwalk interface provides a click and highlight visualisation for colouring points in an embedding according to their distance
in the original high-dimensional space [@Ovchinnikova2019-gf]. The work by 
@Pezzotti2017-cz provides a user guided and modified form of the t-SNE algorithm, that allows users to modified optimisation parameters in real-time.
Similarly, the embedding projector is a web interface to running UMAP, t-SNE
or PCA live in the browser and provides interactions to color points, and
highlights nearest neighbors [@Smilkov2016-hp].

There is no one-size fits all: finding an appropriate embedding for a given
dataset is a difficult and somewhat poorly defined problem. For non-linear methods, there are a lot of parameters to explore that can have an effeft on the resulting visualisation and interpretation. Interfaces for evaluating embeddings
require interaction but should also be able to be incorporated into an analysts workflow. We propose a more pragmatic workflow inspired by incorporating
interactive graphics and tours with embeddings that allows users to see a
global overview of their high dimensional data and assists them with cluster
orientation tasks. This workflow is incorporated into an R package
called `liminal` (Available: https://github.com/sa-lee/liminal).
